Graph Databases for Use with Timeseries IoT Datasets

### 2023-12-04 paper.tex
More changes to final report
More changes to final report
pushed by nathaniel westlake

@@

With the experiments done and the data aquired there were some interesing results that were discovered. One such
discovery was that, within the dataset being used, there were clear sinks and sources for flows as shown in figures
\ref{fig:flowfrom} and \ref{fig:flowto}.

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure1.png}
    \centering
    \caption{The number of flows that were sent from particular IP addresses}
    \label{fig:flowfrom}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure2.png}
    \centering
    \caption{The number of flows that were sent to particular IP addresses}
    \label{fig:flowto}
\end{figure}

A clear pattern of sending and receiving was being established in the data which is a potential indicator of some
sort of attack whether DDoS, Network Scanning or otherwise. To determine this further, the average number of flows
per second was determined as shown in figures \ref{fig:sentfrompersec} and \ref{fig:senttopersec}.

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure3.png}
    \centering
    \caption{The number of flows that were sent from particular IP addresses per second}
    \label{fig:sentfrompersec}
\end{figure}

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure4.png}
    \centering
    \caption{The number of flows that were sent to particular IP addresses per second}
    \label{fig:senttopersec}
\end{figure}

As with figures \ref{fig:flowfrom} and \ref{fig:flowto}, these figures also showed clear trends with definitive
sinks and sources, however this was not enough to fully determine what was going on with this data. To make a more
definitive conclusion the number of ports each IP address contacted was found as shown in figure \ref{fig:portscontacted}.

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure5.png}
    \centering
    \caption{The number of ports contacted per IP address}
    \label{fig:portscontacted}
\end{figure}

This figure provided a more definitive idea about what was going on for some of the IP Addresses. In particular,
192.168.100.148 contacted the most ports likely indicating that it was performing a network scan on one of the
other networks. Similarily could be said for 192.168.100.149 and 192.168.100.150 with less confidence.

@@

\begin{figure}[h]
    \includegraphics[width=8cm]{Figure6.png}
    \centering
    \caption{The number of flows from unique IP addresses coming into a particular IP address}
    \label{fig:incomingflows}
\end{figure}

@@

Clearly the work done here is just the beginning, however the results already gained show great promise of revealing
important information. From so little data and simple queries we can already determine whether two important attacks
occur. However the results found by these queries on this dataset are limited. For one, only simple queries were
and more advanced queries could and should be used in the future in order to solve more complicated problems and
answer more complicated questions. Secondly, the fraction of the dataset used only involved 1,000,000 flows. In
order to get a more accurate picture of a network, millions more flows so be gathered and analyzed. Luckily,
due to the speed of graph databases, this shouldn't be a problem for similar experiments in the future.

@@

\section{Conclusion}
Graph databases proved to be a useful tool for IoT device timeseries data. It is both fast to query and the
information gained provide meaningful answers to interesting questions.

>>>>

The hope of this paper is to find if graph databases are truly a good solution for these users and see
if they are able to answer questions that these users would want to ask about their networks. For instance,
we explore finding whether or not certain attacks had taken place on these systems, a question that would
likely interest anyone running a network of IoT devices.

@@

With the experiments done and the data aquired there were some interesing results that were discovered. One such
discovery was that, within the dataset being used, there were clear sinks and sources for flows as shown in figures
\ref{fig:flowfrom} and \ref{fig:flowto}. These were found by writing a query to determine how many relations were 
being sent both from a node (figure \ref{fig:flowfrom}) and to a node (figure \ref{fig:flowto}).

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure1.png}
    \centering
    \caption{The number of flows that were sent from particular IP addresses}
    \label{fig:flowfrom}
\end{figure}

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure2.png}
    \centering
    \caption{The number of flows that were sent to particular IP addresses}
    \label{fig:flowto}
\end{figure}

A clear pattern of sending and receiving was being established in the data which is a potential indicator of some
sort of attack whether DDoS, Network Scanning or otherwise. To determine this further, the average number of flows
per second was determined as shown in figures \ref{fig:sentfrompersec} and \ref{fig:senttopersec}. This was gained in
a similar fashion to figure \ref{fig:flowfrom} and \ref{fig:flowto} with dividing them further based on their Unix
time stamp and making an average.

While it is good to see how much data was being sent overall, it doesn't really tell very much about what's happening.
In order to determine something like a DDoS attack, the average packets sent pers second was determined to be more useful.

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure3.png}
    \centering
    \caption{The number of flows that were sent from particular IP addresses per second}
    \label{fig:sentfrompersec}
\end{figure}

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure4.png}
    \centering
    \caption{The number of flows that were sent to particular IP addresses per second}
    \label{fig:senttopersec}
\end{figure}

As with figures \ref{fig:flowfrom} and \ref{fig:flowto}, these figures also showed clear trends with definitive
sinks and sources, however this was not enough to fully determine what was going on with this data. To make a more
definitive conclusion the number of ports each IP address contacted was found as shown in figure 
\ref{fig:portscontacted}. This query was performed by counting the unique ports contacted by each IP address.

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure5.png}
    \centering
    \caption{The number of ports contacted per IP address}
    \label{fig:portscontacted}
\end{figure}

This figure provided a more definitive idea about what was going on for some of the IP Addresses. In particular,
192.168.100.148 contacted the most ports likely indicating that it was performing a network scan on one of the
other networks. Similarily could be said for 192.168.100.149 and 192.168.100.150 with less confidence. This query
was performed by looking at the unique nodes that connected to the IP addresses.

@@

\begin{figure}[htbp]
    \includegraphics[width=8cm]{Figure6.png}
    \centering
    \caption{The number of flows from unique IP addresses coming into a particular IP address}
    \label{fig:incomingflows}
\end{figure}

@@

Clearly the work done here is just the beginning, however the results already gained show great promise of revealing
important information. From so little data and simple queries we can already determine whether two important attacks
occur. From data such as flows per second and unique devices connecting it is possible to figure out if a DDoS attack
is occuring. From data such as ports contacted it is possible to see if the network is being scanned.
However the results found by these queries on this dataset are limited. For one, only simple queries were
and more advanced queries could and should be used in the future in order to solve more complicated problems and
answer more complicated questions. Secondly, the fraction of the dataset used only involved 1,000,000 flows. In
order to get a more accurate picture of a network, millions more flows so be gathered and analyzed. Luckily,
due to the speed of graph databases, this shouldn't be a problem for similar experiments in the future.

@@

\section{Conclusion}
While there are many avenues this work could take in the future, the goal that this paper set out with was
achieved. It was shown that graph databases could effectively use IoT timeseries data to answer questions
that would be of interest to both home and small network users. It has been shown that graph databases work
on timeseries databases with relatively simple queries to produce results for real world applications
